---
title: "Random Forest"
author: "Sixtus Dakurah"
date: "3/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```


```{r}
library(Metrics)
```


```{r}
# expose the modules here
packages.Self <- modules::use("libs.R")
graphics.Self <- modules::use("graphics.R")
munging.Self <- modules::use("mungin.R")
network.Self <- modules::use("network.R")
```

```{r}
# load libraries
packages.Self$getPackages()
```


```{r message=FALSE, warning=FALSE}
movie_summary_net <- network.Self$getDataFile("https://sixtusdakurah.com/projects/box-office/movie_summary.csv")
movie_ratings <- network.Self$getDataFile("https://sixtusdakurah.com/projects/box-office/movie_ratings.csv")
#weekend_boxoffice <- network.Self$getDataFile("https://sixtusdakurah.com/projects/box-office/weekend_boxoffice.csv")
```


```{r}
dim(movie_summary_net)
dim(movie_ratings)
```


```{r}
#movie_summary_net$international_box_office <- as.numeric(movie_summary_net$international_box_office)
movie_summary_net$inflation_adjusted_domestic_box_office <- as.numeric(levels(movie_summary_net$inflation_adjusted_domestic_box_office))[movie_summary_net$inflation_adjusted_domestic_box_office]
head(movie_summary_net)
```

```{r}
dim(movie_summary_net[is.na(movie_summary_net$inflation_adjusted_domestic_box_office), ])
```


```{r}
movie_merged <- merge(movie_summary_net, movie_ratings, by = "movie_odid")
```


```{r}
glimpse(movie_merged)
```



```{r}
head(movie_merged)
```


## Repeat of EDA - Continuous Features

# remove zero opening weekedn revenue

```{r}
movie_merged <- movie_merged[movie_merged$opening_weekend_revenue>0, ]
```


## add the inflation factor and adjust the domestic box office ... just using the inflation factor doesn't seem right. 
## The opening weekend revenue is just for a week, the domestic boxoffice is for a whole year

```{r}
movie_merged$inflation_factor <- movie_merged$inflation_adjusted_domestic_box_office/movie_merged$domestic_box_office
# now adjust the opening weekend revenue for inflation
movie_merged$opening_weekend_revenue <- movie_merged$opening_weekend_revenue*movie_merged$inflation_factor

head(movie_merged)
```


```{r}
# add the reciprocal transform
#movie_merged$rev_recp <- 1/(movie_merged$opening_weekend_revenue)
ggplot(movie_merged[movie_merged$opening_weekend_revenue> median(movie_merged$opening_weekend_revenue), ], aes(log10(opening_weekend_revenue))) + geom_histogram() # + xlab("Production Budget") + ylab("Opening Weekend Revenue")
```



```{r}
ggplot(movie_merged[log10(movie_merged$opening_weekend_revenue)>6.25, ], aes(log10(opening_weekend_revenue))) + geom_histogram() # + xlab("Production Budget") + ylab("Opening Weekend Revenue")
```

## put that to a dummy dataset

```{r}
upper.data.all <- movie_merged[movie_merged$opening_weekend_revenue > median(movie_merged$opening_weekend_revenue)&&log10(movie_merged$opening_weekend_revenue)>6.25, ]
ggplot(movie_merged[movie_merged$opening_weekend_revenue > median(movie_merged$opening_weekend_revenue)&&log10(movie_merged$opening_weekend_revenue)>6.25, ], aes(log10(opening_weekend_revenue))) + geom_histogram() 
```



```{r}
dim(movie_merged[movie_merged$opening_weekend_revenue==0, ])
```

```{r}
dim(movie_merged[is.na(movie_merged$inflation_adjusted_domestic_box_office), ])
```



```{r}
#ggplot(movie_merged, aes(x = production_budget, y=opening_weekend_revenue)) + geom_point() + xlab("Production Budget") + ylab("Opening Weekend Revenue")
ggplot(movie_merged, aes(x = production_budget, y=opening_weekend_revenue)) + geom_point() + xlab("Production Budget") + ylab("Opening Weekend Revenue")
```


## convert run time to continuous

```{r}
movie_merged$running_time <- as.numeric(movie_merged$running_time)
```

```{r}
ggplot(movie_merged, aes(x = running_time, y=opening_weekend_revenue)) + geom_point() + xlab("Running Time") + ylab("Opening Weekend Revenue")
```

## there is a sharp divide in the runtime for various movies. but they does not appear to be a strong association between the two



##  A look at other variables

```{r}
movie_merged$distributor <- fct_explicit_na(movie_merged$distributor, na_level = "Unknown")

movie_merged %>%
   group_by(distributor) %>%
   summarise(movie_count = n()) %>%
   arrange(desc(movie_count)) %>%
   filter(!is.na(distributor))
```


## Craete a top distributor category for distributors with categories above 2999

```{r}
movie_merged$top_distributor[movie_merged$distributor=='Warner Bros.'] <- 'top distributor'
movie_merged$top_distributor[movie_merged$distributor=='Sony Pictures'] <- 'top distributor'
movie_merged$top_distributor[movie_merged$distributor=='Lionsgate'] <- 'top distributor'
movie_merged$top_distributor[movie_merged$distributor=='IFC Films'] <- 'top distributor'

movie_merged$top_distributor[movie_merged$distributor=='Sony Pictures Classics'] <- 'top distributor'
movie_merged$top_distributor[movie_merged$distributor=='Walt Disney'] <- 'top distributor'
movie_merged$top_distributor[movie_merged$distributor=='20th Century Fox'] <- 'top distributor'
movie_merged$top_distributor[movie_merged$distributor=='Paramount Pictures'] <- 'top distributor'
movie_merged$top_distributor[movie_merged$distributor=='Universal'] <- 'top distributor'
#movie_merged$top_distributor[movie_merged$distributor=='Miramax'] <- 'medium distributor'
#movie_merged$top_distributor[movie_merged$distributor=='Magnolia Pictures'] <- 'medium distributor'


# lump the null category to the rest
movie_merged$top_distributor[is.na(movie_merged$top_distributor)] <- 'other distributor'

```

## look at it's effect on revenue

```{r}
ggplot(movie_merged, aes(y =opening_weekend_revenue, x=top_distributor)) + geom_boxplot() + xlab("Top Distributor") + ylab("Opening Weekend Revenue")
```


```{r}
ggplot(movie_merged, aes(y = production_budget , x=top_distributor)) + geom_boxplot() + xlab("Top Distributor") + ylab("Production Budget")
```


## use this to create the production company size

```{r}
movie_merged$production_comp_size[movie_merged$top_distributor=='top distributor'] <- 'big producer'
# lump other category to the rest
movie_merged$production_comp_size[is.na(movie_merged$production_comp_size)]  <- 'small producer'
```


```{r}
ggplot(movie_merged, aes(y =opening_weekend_revenue , x=production_comp_size)) + geom_boxplot() + xlab("Production Company Size") + ylab("Opening Weekend Revenue")
```



```{r}
ggplot(movie_merged, aes(y =opening_weekend_revenue , x=production_comp_size)) + geom_bar(stat = "summary", fun.y="mean") + xlab("Production Company Size") + ylab("Opening Weekend Revenue")
```


```{r}
movie_merged$release_date_char <- as.character(movie_merged$release_date)
movie_merged$release_date_mod <- parse_date_time(as.character(movie_merged$release_date_char), "ymd")


movie_merged$year_released <- ymd(movie_merged$release_date_mod) %>% 
   lubridate::year()  # Grab year.

movie_merged$quarter_released <- ymd(movie_merged$release_date_mod) %>%
   lubridate::quarter()  # Grab quarter.

movie_merged$month_released <- ymd(movie_merged$release_date_mod) %>% 
   lubridate::month(label = TRUE, abbr = FALSE)  # Grab month.

movie_merged$week_released <- ymd(movie_merged$release_date_mod) %>%
   lubridate::week()  # Grab week.

movie_merged$weekday_released <- ymd(movie_merged$release_date_mod) %>%
   lubridate::wday(label = TRUE, abbr = FALSE)  # Grab weekday.

```


```{r}
head(movie_merged)
```


## look at the median revenue by quarter

```{r}
ggplot(movie_merged, aes(y =opening_weekend_revenue , x=quarter_released)) + geom_bar(stat = "summary", fun.y="median") + xlab("Release Quater") + ylab("Opening Weekend Revenue")
```



## look at the median revenue by quarter

```{r}
ggplot(movie_merged, aes(y =opening_weekend_revenue , x=month_released)) + geom_bar(stat = "summary", fun.y="median") + xlab("Release Month") + ylab("Opening Weekend Revenue") + coord_flip()
```



## look at the median revenue by genre

```{r}
movie_merged$genre <- fct_explicit_na(movie_merged$genre, na_level = "Unknown")
ggplot(movie_merged, aes(x=genre)) + geom_bar(position = "dodge") + xlab("Genre") + ylab("Count") + coord_flip()
```

```{r}
movie_merged$genre <- fct_explicit_na(movie_merged$genre, na_level = "Unknown")
ggplot(movie_merged, aes(x=genre, y=opening_weekend_revenue)) + geom_bar(position = "dodge", stat="summary", fun.y="median") + xlab("Genre") + ylab("Median Opening Weekend Revenue") + coord_flip()
```

## create a new category to take care of that

```{r}
movie_merged$top_grossing_genre[movie_merged$genre=='Adventure'] <- 'top genre'
movie_merged$top_grossing_genre[movie_merged$genre=='Action'] <- 'top genre'
movie_merged$top_grossing_genre[movie_merged$genre=='Horror'] <- 'medium genre'
#movie_merged$top_grossing_genre[movie_merged$genre=='Thriller/Suspense'] <- 'medium genre'
#movie_merged$top_grossing_genre[movie_merged$genre=='Romantic comedy'] <- 'medium genre'

# lump the null category to the rest
movie_merged$top_grossing_genre[is.na(movie_merged$top_grossing_genre)] <- 'other genre'
```

```{r}
ggplot(movie_merged, aes(x=top_grossing_genre, y=opening_weekend_revenue)) + geom_bar(position = "dodge", stat="summary", fun.y="median") + xlab("Genre") + ylab("Median Opening Weekend Revenue") + coord_flip()
```


## A loot at rating

```{r}
movie_merged$rating <- fct_explicit_na(movie_merged$rating, na_level = "Unknown")
ggplot(movie_merged, aes(x=rating)) + geom_bar(position = "dodge") + xlab("Rating") + ylab("Count") + coord_flip()
```


```{r}
ggplot(movie_merged, aes(x=rating, y=opening_weekend_revenue)) + geom_bar(position = "dodge", stat="summary", fun.y="median") + xlab("Rating") + ylab("Median Opening Weekend Revenue") + coord_flip()
```


## create a new variable for key words count in title and title length

```{r}
movie_merged$title_length <- str_length(movie_merged$display_name.x)
movie_merged$description_length <- str_length(movie_merged$reason)
```


```{r}
ggplot(movie_merged, aes(x=title_length, y=opening_weekend_revenue)) + geom_point() + xlab("Title Length") + ylab("Opening Weekend Revenue")
```

```{r}
ggplot(movie_merged, aes(x=description_length, y=opening_weekend_revenue)) + geom_point() + xlab("Description Length") + ylab("Opening Weekend Revenue")

```


```{r}
movie_merged$source<- fct_explicit_na(movie_merged$source, na_level = "Unknown")
movie_merged$creative_type<- fct_explicit_na(movie_merged$creative_type, na_level = "Unknown")
movie_merged$production_method<- fct_explicit_na(movie_merged$production_method, na_level = "Unknown")
```


```{r}
merged_subset <- subset(movie_merged,
                        select = c(
                          opening_weekend_revenue, 
                          running_time, 
                          sequel, 
                          creative_type, 
                          source, 
                          genre, 
                          production_method,
                          production_budget, 
                          rating, 
                          top_distributor, 
                          production_comp_size, 
                          quarter_released, 
                          month_released, 
                          year_released,
                          top_grossing_genre, 
                          title_length, 
                          description_length))
```


```{r}
aggr(merged_subset, sortVars = TRUE, prop = FALSE, cex.axis = .35, 
     numbers = TRUE, col = c('grey99','red'))
```

## impute categorical missing with the mean

```{r}
merged_subset$description_length[is.na(merged_subset$description_length)] <- mean(merged_subset$description_length, na.rm = TRUE)
```


```{r}
aggr(merged_subset, sortVars = TRUE, prop = FALSE, cex.axis = .35, 
     numbers = TRUE, col = c('grey99','red'))
```


## log transfrom skewed data

```{r}
merged_subset <- mutate(merged_subset,
                           production_budget = log10(production_budget + 1),
                           #opening_weekend_revenue = log10(opening_weekend_revenue),
                        #opening_weekend_revenue = rev_recp,
                        running_time = log10(running_time + 1),
                        title_length  = log10(title_length + 1),
                        description_length = log10(description_length + 1)
                        
                        )
```

```{r}
head(merged_subset)
```

## create a linear model to predict for 0 budget

```{r}
# Create linear model to predict budget. 
lm_budget <- lm(production_budget ~ running_time + sequel + production_method +
                   top_grossing_genre + production_comp_size + quarter_released , 
                data = merged_subset[merged_subset$production_budget!=0, ])
summary(lm_budget)
```

```{r}
# Predict all 0 in budget with lm_budget.
merged_subset$production_budget[merged_subset$production_budget==0] <- predict(lm_budget)
```

```{r}
merged_subset$budget_year_ratio <- merged_subset$production_budget/merged_subset$year_released
```

```{r}
merged_subset <-merged_subset %>% mutate_if(is.character, as.factor)
merged_subset$month_released <- factor(merged_subset$month_released, ordered = FALSE)
merged_subset$quarter_released <- factor(merged_subset$quarter_released)
```


```{r}
summary(merged_subset$opening_weekend_revenue)
```

```{r}
summary( (merged_subset[merged_subset$opening_weekend_revenue >= 8000 & merged_subset$opening_weekend_revenue<40000, ])$opening_weekend_revenue)
```


```{r}
par(mfrow=c(1,3))
ggplot(merged_subset, aes(opening_weekend_revenue)) + geom_histogram()

ggplot(merged_subset[merged_subset$opening_weekend_revenue < 8000, ], aes(opening_weekend_revenue)) + geom_histogram()

ggplot(merged_subset[merged_subset$opening_weekend_revenue >= 8000 & merged_subset$opening_weekend_revenue<11552, ], aes(opening_weekend_revenue)) + geom_histogram() 

ggplot(merged_subset[merged_subset$opening_weekend_revenue >= 11552 & merged_subset$opening_weekend_revenue<16928, ], aes(opening_weekend_revenue)) + geom_histogram() 

dim(merged_subset[merged_subset$opening_weekend_revenue < 6.25, ])
dim(merged_subset[merged_subset$opening_weekend_revenue >= 6.25 && merged_subset$opening_weekend_revenue<8, ])
```

```{r}
summary(merged_subset[merged_subset$opening_weekend_revenue < 6.25, ])
summary(merged_subset[merged_subset$opening_weekend_revenue >= 6.25 & merged_subset$opening_weekend_revenue >= 6.25, ])
```




```{r}
merged_subset2 <- merged_subset[merged_subset$opening_weekend_revenue >= 6.25, ]
merged_subset3 <- merged_subset[merged_subset$opening_weekend_revenue < 6.25, ]
```


## train test split

```{r}
ggplot(merged_subset2, aes((opening_weekend_revenue))) + geom_histogram()
#merged_subset <- merged_subset[merged_subset$opening_weekend_revenue<=6.25, ]
```



```{r}
train <- merged_subset[merged_subset$year_released!='2019', ]
dim(train)
test <- merged_subset[merged_subset$year_released=='2019', ]
dim(test)
```

```{r}
head(train)
```


```{r}
set.seed(222)

rf_model <- randomForest(opening_weekend_revenue ~ .,
                         data = train, 
                         ntree = 10001,
                         replace = TRUE,
                         nodesize = 5,
                         importance = TRUE); 
print(rf_model)
```


```{r}
# Create an object for importance of variables
importance <- importance(rf_model) 

# Create data frame using importance. 
varImportance <- data.frame(Variables = row.names(importance), 
                            Importance = round(importance[,'IncNodePurity'], 0))

# Create interactive plot.  
ggplotly(ggplot(varImportance, aes(x = reorder(Variables, Importance), 
                           y = Importance, fill = Importance)) +
       geom_bar(stat='identity') + 
       labs(title = 'Importance of predictors', x = 'Predictors', y = 'rmsle') +
       coord_flip() + 
       theme_light())
```


```{r}
prediction <- predict(rf_model, test)
```


```{r}
#id <- (movie_merged[movie_merged$year_released=='2019', ])$movie_odid
solution <- data.frame(revenue_pred_ls = prediction)
solution$revenue_ls <- test$opening_weekend_revenue
#solution <- as_tibble(solution) %>% mutate(revenue_pred = 10^revenue_pred_ls)
#solution <- as_tibble(solution) %>% mutate(revenue = 10^revenue_ls)
```

```{r}
(solution)
```

```{r}
sqrt(mean((solution$revenue_pred_ls -solution$revenue_ls)**2))
```




```{r}
train1 <- merged_subset3[merged_subset2$year_released!='2019', ]
dim(train1)
test1 <- merged_subset3[merged_subset2$year_released=='2019', ]
dim(test1)
```



```




