---
title: "Regression Analysis"
author: "Sixtus Dakurah"
date: "2/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# remove env var to start with
rm(list=ls())
# mandatory install
library(modules)
# set wkdir
setwd("")
library(stargazer)
```

```{r}
library(Metrics)
```



## load modules

```{r}
# expose the modules here
packages.Self <- modules::use("libs.R")
graphics.Self <- modules::use("graphics.R")
munging.Self <- modules::use("mungin.R")
network.Self <- modules::use("network.R")
```


## get data from network

```{r message=FALSE, warning=FALSE}
movie_summary_net <- network.Self$getDataFile("https://sixtusdakurah.com/projects/box-office/movie_summary.csv")
movie_ratings <- network.Self$getDataFile("https://sixtusdakurah.com/projects/box-office/movie_ratings.csv")
#weekend_boxoffice <- network$getDataFile("https://sixtusdakurah.com/projects/box-office/weekend_boxoffice.csv")
```


# take a head look

```{r}
head(movie_summary_net)
```


```{r}
dim(movie_summary_net)
dim(movie_ratings)
```


# merge the ratings to the movie summary

```{r}
movie_summary <- merge(movie_summary_net, movie_ratings,  by = "movie_odid")
head(movie_summary)
```

# take a look at the production weekend revenue

```{r}
summary(movie_summary$opening_weekend_revenue)
```

```{r}
dim(movie_summary[movie_summary$opening_weekend_revenue==0, ])
```

# about 3000 movies did not have opening weekend revenue

```{r}
# less the median
dim(movie_summary[movie_summary$opening_weekend_revenue < median(movie_summary$opening_weekend_revenue), ])
```

# working with movies about the median

```{r}
movie_summary_above_median <- movie_summary[movie_summary$opening_weekend_revenue >= median(movie_summary$opening_weekend_revenue), ]
```

```{r}
summary(movie_summary_above_median$opening_weekend_revenue)
```

# take a look at the distribution

```{r}
par(mfrow=c(1,2))
hist(movie_summary_above_median$opening_weekend_revenue, main = "Original", xlab = "OPW Revenue")
hist(log(movie_summary_above_median$opening_weekend_revenue), main = "Log transformed", xlab = "Log OPW Revenue")
```

# original heavily skewed to the right and the log transform is bimodal! Model the log of the opening weekend might be good!

# taking a look at the production budget

```{r}
summary(movie_summary_above_median$production_budget)
```

# some prod budgets are zero!

```{r}
dim(movie_summary[movie_summary_above_median$production_budget==0, ])
```

```{r}
summary( (movie_summary_above_median[movie_summary_above_median$production_budget>0, ])$production_budget )
```

```{r}
hist(log((movie_summary_above_median[movie_summary_above_median$production_budget>1000000, ])$production_budget ), main = "Production Budget", xlab = "Budget")
```

# only look at movies with production budget of atleast a million dollar

```{r}
movie_summary_prod_ <- movie_summary_above_median[movie_summary_above_median$production_budget>1000000, ]
dim(movie_summary_prod_)
```

```{r}
par(mfrow=c(1,2))
hist(movie_summary_prod_$opening_weekend_revenue, main = "Original", xlab = "OPW Revenue")
hist(log(movie_summary_prod_$opening_weekend_revenue), main = "Log transformed", xlab = "Log OPW Revenue")
```


# look at the scatter plot of weekend revenue and production budget

```{r}
plot( x =log(movie_summary_prod_$production_budget), y= movie_summary_prod_$opening_weekend_revenue, xlab = "Budget", ylab = "Log revenue")
```

# add the log revenue var

```{r}
movie_summary_prod_$log_revenue <- log(movie_summary_prod_$opening_weekend_revenue)
```


# test regression model


```{r}
test.lm.model <- lm(log_revenue~production_budget, data = movie_summary_prod_)
summary(test.lm.model)
```

```{r}
sum(exp(test.lm.model$residuals))
```



# look at run time

```{r}
movie_summary_prod_$running_time <- as.numeric(movie_summary_prod_$running_time)
summary(movie_summary_prod_$running_time)
```

```{r}
hist(movie_summary_prod_[movie_summary_prod_$running_time > 100, ]$running_time)
```

```{r}
dim(movie_summary_prod_[movie_summary_prod_$running_time <= 100, ])
```

# modeling the runtime directly might not be useful put into categories

```{r}
movie_summary_prod_$running_length <- sapply(movie_summary_prod_$running_time, function(t){(if (t < 100 ) '1' else '0')})
```

```{r}
table((movie_summary_prod_$running_length))
```

```{r}
test.lm.model <- lm(log_revenue~production_budget + running_time:running_length, data = movie_summary_prod_)
summary(test.lm.model)
```

```{r}
sum(exp(test.lm.model$residuals))
```


## look at sequel

```{r}
movie_summary_prod_ <- (movie_summary_prod_[movie_summary_prod_$sequel!='NULL', ])
dim(movie_summary_prod_)
```


```{r}
test.lm.model <- lm(log_revenue~production_budget + running_time:running_length + sequel, data = movie_summary_prod_)
summary(test.lm.model)
sum(exp(test.lm.model$residuals))
```


## take a  look at the release date

```{r}
# get the release year and month
movie_summary_prod_$release_date<- as.Date(movie_summary_prod_$release_date)
movie_summary_prod_$release_year <- (format(movie_summary_prod_$release_date, '%Y'))
movie_summary_prod_$release_month <- ((format(movie_summary_prod_$release_date, '%m')))
```

```{r}
table(movie_summary_prod_$release_month)
```

```{r}
library(ggplot2)
ggplot(movie_summary_prod_, aes(as.factor(release_month), opening_weekend_revenue)) + 
    geom_boxplot() + 
    labs(y = "Percentage correct", x = "Categorical variable");
```

### its like may june august and november

```{r}
movie_summary_prod_$season <- sapply(movie_summary_prod_$release_month, function(t){(if (t == 05 ) '1' else if (t==06) {'1'} else if (t==07){'1'} else if(t==11){'1'} else  '0')})
```

```{r}
table(movie_summary_prod_$season)
```

```{r}
test.lm.model <- lm(log_revenue~production_budget + running_time:running_length + sequel + release_month, data = movie_summary_prod_)
summary(test.lm.model)
sum(exp(test.lm.model$residuals))
```


## look at opening weekend theaters

```{r}
summary(movie_summary_prod_$opening_weekend_theaters)
```

```{r}
movie_summary_th_ <- (movie_summary_prod_[movie_summary_prod_$opening_weekend_theaters > 0  , ])
```

```{r}
summary(movie_summary_th_$opening_weekend_theaters)
```

```{r}
hist(movie_summary_th_$opening_weekend_theaters)
```

```{r}
test.lm.model <- lm(log_revenue~production_budget + sequel + running_time + season + opening_weekend_theaters, data = movie_summary_th_)
summary(test.lm.model)
print(paste( "RMSE; ",  sqrt(mean(( ( exp(test.lm.model$fitted.values) * mean(exp(test.lm.model$residuals)) ) - movie_summary_th_$opening_weekend_revenue)**2))))
```

## this is loooking pretty good!

# take a look at genre -- remove null



```{r}
table(movie_summary_th_$genre)
```

```{r}
movie_summary_gen <- movie_summary_th_[movie_summary_th_$genre!='NULL', ]
dim(movie_summary_gen)
movie_summary_gen$genre <- as.character(movie_summary_gen$genre)
movie_summary_gen <- (movie_summary_gen[!is.na(movie_summary_gen$genre), ])
table(movie_summary_gen$genre)
```

```{r}
unique(movie_summary_gen$genre)
```




```{r}
ggplot(movie_summary_gen, aes(as.factor(genre), opening_weekend_revenue)) + 
    geom_boxplot() + 
    labs(y = "Percentage correct", x = "Categorical variable");
```

```{r}
hist(movie_summary_gen[movie_summary_gen$opening_weekend_revenue < 19019731 & movie_summary_gen$opening_weekend_revenue >2000000, ]$opening_weekend_revenue)
```



```{r}
movie_summary_gen$gen_cat <- sapply(as.character(movie_summary_gen$genre), function(t){(if (t == "Adventure" ) '1' else if(t=="Action") {'1'} else '0' )})
```

```{r}
table(movie_summary_gen$gen_cat)
```

```{r}
test.lm.model <- lm(opening_weekend_revenue ~ opening_weekend_theaters, data = movie_summary_gen)
summary(test.lm.model)
print(paste( "RMSE; ",  sqrt(mean(( ( exp(test.lm.model$fitted.values) * mean(exp(test.lm.model$residuals)) ) - movie_summary_gen$opening_weekend_revenue)**2))))
```



```{r}
plot(test.lm.model$residuals)
```

```{r}
movie_summary_gen$release_year <- as.numeric(movie_summary_gen$release_year)
movie_summary_gen$season <- factor(movie_summary_gen$season, levels = unique(movie_summary_gen$season))
movie_summary_gen$genre <- factor(movie_summary_gen$genre, levels = unique(movie_summary_gen$genre))
head(movie_summary_gen)
```


```{r}
#divide into train and test set.seed()
train <- movie_summary_gen[movie_summary_gen$release_year < 2019, ]
test <- movie_summary_gen[movie_summary_gen$release_year == 2019, ]
```

```{r}
table(test$release_year)
table(train$release_year)
```


```{r}
test.lm.model <- lm(log_revenue ~ + opening_weekend_theaters + sequel + running_time + season + genre + creative_type + source + production_method, data = train)
summary(test.lm.model)
print(paste( "RMSE; ",  sqrt(mean(( ( exp(test.lm.model$fitted.values) * mean(exp(test.lm.model$residuals)) ) - train$opening_weekend_revenue)**2))))
```


## Construction a new datapoint



```{r}
# Onward
sequel <- '0'
running_time <- 114
season <- '0'
opening_weekend_theaters <- 4529
genre <- 'Adventure' 
creative_type <- 'Dramatization'
source <- 'Original Screenplay'  
production_method <- 'Digital Animation'
#log_revenue <- log(revenue)
newdf <- data.frame(revenue, sequel, running_time, season, opening_weekend_theaters, genre, creative_type, source, production_method)

newdf
```

```{r}
yhat <- predict.lm(test.lm.model, newdf)
yhat
```

```{r}
head(test)
```


```{r}
exp(yhat)
```



```{r}
# Theway Back
sequel <- '0'
running_time <- 108
season <- '0'
opening_weekend_theaters <- 3014
genre <- 'Drama' 
creative_type <- 'Contemporary Fiction'
source <- 'Original Screenplay'  
production_method <- 'Live Action'
#log_revenue <- log(revenue)
newdf1 <- data.frame(revenue, sequel, running_time, season, opening_weekend_theaters, genre, creative_type, source, production_method)

newdf1
```

```{r}
predict.lm(test.lm.model, newdf1)
exp(predict.lm(test.lm.model, newdf1))
```


```{r}

# Onward
sequel <- '0'
running_time <- 119
season <- '0'
opening_weekend_theaters <- 2286
genre <- 'Drama' 
creative_type <- 'Historical Fiction'
source <- 'Based on Fiction Book/Short Story'  
production_method <- 'Live Action'
#log_revenue <- log(revenue)
newdf2 <- data.frame(revenue, sequel, running_time, season, opening_weekend_theaters, genre, creative_type, source, production_method)

newdf2

```

```{r}
predict.lm(test.lm.model, newdf2)
exp(predict.lm(test.lm.model, newdf2))
```

### Trying GBM

```{r}
# model
library(gbm)
set.seed(123)
M <- gbm::gbm( opening_weekend_revenue ~ opening_weekend_theaters, 
             # the formula for the model (recall that the period means, "all 
             # variables in the data set")
             data = train, 
             # data set
             #distribution = 'tdist',
             # i'm using tdist since it can accomodate heavire tails
             shrinkage = 0.01, 
             # a shrinkage parameter applied to each tree in the expansion. 
             # Also known as the learning rate or step-size reduction; 0.001 
             # to 0.1 usually work, but a smaller learning rate typically 
             # requires more trees.
             interaction.depth = 3, 
             # Integer specifying the maximum depth of each tree (i.e., the 
             # highest level of variable interactions allowed). A value of 1 
             # implies an additive model, a value of 2 implies a model with up
             #  to 2-way interactions
             n.minobsinnode = 5,
             # Integer specifying the minimum number of observations in the 
             # terminal nodes of the trees. Note that this is the actual number 
             # of observations, not the total weight.
             n.trees = 5000 
             # Integer specifying the total number of trees to fit. This is 
             # equivalent to the number of iterations and the number of basis 
             # functions in the additive expansion.
             )
```


```{r}
perf_gbm_ = gbm.perf(M, method = "OOB")
```

```{r}
movie_prediction_ <- stats::predict(
                           # the model from above
                          object = M, 
                          # the testing data
                          newdata = newdf,
                          # this is the number we calculated above
                          n.trees = perf_gbm_)
rmse_fit_ <- Metrics::rmse(actual = test$opening_weekend_revenue, 
                           predicted = movie_prediction_)
print(rmse_fit_)
```

```{r}
movie_prediction_
```



```{r}
movie_prediction_ <- stats::predict(
                           # the model from above
                          object = M, 
                          # the testing data
                          newdata = newdf1,
                          # this is the number we calculated above
                          n.trees = perf_gbm_)
rmse_fit_ <- Metrics::rmse(actual = test$opening_weekend_revenue, 
                           predicted = movie_prediction_)
print(rmse_fit_)
```


```{r}
movie_prediction_ <- stats::predict(
                           # the model from above
                          object = M, 
                          # the testing data
                          newdata = newdf2,
                          # this is the number we calculated above
                          n.trees = perf_gbm_)
rmse_fit_ <- Metrics::rmse(actual = test$opening_weekend_revenue, 
                           predicted = movie_prediction_)
print(rmse_fit_)
```







### look at the inflation data

```{r}
inflation_data <- read.csv("data/cpiai_csv.csv")
```

```{r}
# change to date variable
inflation_data$Date <- as.Date(inflation_data$Date)
inflation_data$year <- (format(inflation_data$Date, '%Y'))
inflation_data$month <- ((format(inflation_data$Date, '%m')))
View(inflation_data)
```


### Munging


```{r}
# convert runtime to integer
movie_summary$running_time <- as.integer(movie_summary$running_time)

# convert inflation adjusted box office from factors to double
movie_summary$inflation_adjusted_domestic_box_office <- as.double(movie_summary$inflation_adjusted_domestic_box_office)

# get the release year and month
movie_summary$release_date<- as.Date(movie_summary$release_date)
movie_summary$release_year <- (format(movie_summary$release_date, '%Y'))
movie_summary$release_month <- ((format(movie_summary$release_date, '%m')))

## drop unnecessary columns

cols.drop <- c("display_name", "od_name", "theatrical_engagements ")
```


```{r}
movie_summary$naive_inflation_factor <- movie_summary$inflation_adjusted_domestic_box_office/movie_summary$domestic_box_office
head(movie_summary)
```



```{r}
# look at the summary statistics
summary(movie_summary)
```


```{r}
# look at the summary of movies without ratings
rated_ids <- movie_ratings$movie_odid
non_rated <- movie_summary_net[!(movie_summary_net$movie_odid %in% rated_ids),]
dim(non_rated)
```

```{r}
head(movie_ratings)
```


```{r}
head(movie_summary_net)
```


```{r}
head(movie_summary)
```


```{r}
summary(non_rated)
```

```{r}
dim(non_rated)
sum(non_rated$opening_weekend_revenue == 0)
```

```{r}
summary(movie_summary)
```



### Subset data 

```{r}
movie_summary_median <- movie_summary[movie_summary$opening_weekend_revenue >= median(movie_summary$opening_weekend_revenue), ]
#head(movie_summary_median)
summary(movie_summary_median) # this does not make any difference
```


######~~~~~~~~~~~~~~~~~~~~~~~~~~
###~~~~~~~~~~~~~~~~~~~~
###  Working with the median data
###~~~~~~~~~~~~~~~~~~~~
######~~~~~~~~~~~~~~~~~~~~~~~~~~


### Build the model

```{r}
movie_summary_median1 <- movie_summary_median[movie_summary_median$production_budget!=0, ]
movie_summary_median1 <- movie_summary_median1[movie_summary_median1$opening_weekend_revenue!=0, ]
# add the log scales
movie_summary_median1$log_rev <- log(movie_summary_median1$opening_weekend_revenue)
movie_summary_median1$log_prod <- log(movie_summary_median1$production_budget)
train.data.m1 <- movie_summary_median1[movie_summary_median1$release_year != "2019", ]
dim(train.data.m1)
test.data.m1 <- movie_summary_median1[movie_summary_median1$release_year == "2019", ]
dim(test.data.m1)
```



```{r}
names(train.data.m)
```


## Initial model

```{r}
library(olsrr)
lm.model.median1 <- lm(log(opening_weekend_revenue) ~ log(production_budget) + running_time +  opening_weekend_theaters + release_month  + source + creative_type + production_method + genre + sequel + rating, data  =  train.data.m1)
#k <- ols_step_all_possible(lm.model.median1)
summary(lm.model.median1)
```

```{r}
#stargazer(lm.model.median1)
```



### performing best subset selection

```{r}
k.best.subset <- ols_step_best_subset(lm.model.median1)
```

```{r}
library(stargazer)
stargazer(k.best.subset)
```


```{r}
k.best.subset
```




```{r}
k[k$adjr==max(k$adjr), ]
```


```{r}
print(paste("Train MSE: ", sqrt(mean(lm.model.median1$residuals)^2)))
#predictions <- predict(lm.model.median1,test.data1.m)
#print(paste("Test MSE: ", sqrt(mean((test.data.m1$opening_weekend_revenue - exp(predictions)) ^ 2))))
```


### Group months into seasons, and ratings into whether it has been rated or not

```{r}
movie_summary_median2 <- movie_summary_median
movie_summary_median2$season<- rep("other", nrow(movie_summary_median2))
movie_summary_median2$is_rated<- rep("yes", nrow(movie_summary_median2))
movie_summary_median2[movie_summary_median2$rating == "Not Rated", ][, "is_rated"] <- "no"
movie_summary_median2[movie_summary_median2$release_month == "05", ][, "season"] <- "summer"
movie_summary_median2[movie_summary_median2$release_month == "06", ][, "season"] <- "summer"
movie_summary_median2[movie_summary_median2$release_month == "07", ][, "season"] <- "summer"
movie_summary_median2[movie_summary_median2$release_month == "12", ][, "season"] <- "yuletide"
head(movie_summary_median2)
```

```{r}
train.data.m2 <- movie_summary_median2[movie_summary_median2$release_year != "2019", ]
dim(train.data.m2)
test.data.m2 <- movie_summary_median2[movie_summary_median2$release_year == "2019", ]
dim(test.data.m2)
```


```{r}
lm.model.median2 <- lm(opening_weekend_revenue ~ production_budget + season + genre + source + running_time + creative_type + production_method + opening_weekend_theaters + sequel + is_rated, data  =  train.data.m2)
summary(lm.model.median2)
```


```{r}
print(paste("Train MSE: ", sqrt(mean(lm.model.median2$residuals)^2)))
predictions1 <- predict(lm.model.median2,test.data.m2)
print(paste("Test MSE: ", sqrt(mean((test.data.m2$opening_weekend_revenue - predictions1) ^ 2))))
```

```{r}
  
```

